---
title: "IPL_Prediction"
author: "Abhineet Ishan Kevin Partha"
date: "4 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


IPL2018 Auction Price Prediction
Three separate models for bowlers, batsman and allrounder players

```{r 1}
library(MASS)
library(glmnet)
setwd("C:/Users/Ishan1/Desktop/IPL_Prediction")
```

# Reading data
```{r 2}
batting2014.df<-read.csv('batsman_2014.csv', header=T)
batting2015.df<-read.csv('batsman_2015.csv', header=T)
batting2016.df<-read.csv('batsman_2016.csv', header=T)
batting2017.df<-read.csv('batsman_2017.csv', header=T)


bowling2014.df<-read.csv('baller_2014.csv', header=T)
bowling2015.df<-read.csv('baller_2015.csv', header=T)
bowling2016.df<-read.csv('baller_2016.csv', header=T)
bowling2017.df<-read.csv('baller_2017.csv', header=T)


allrounder2014.df<-read.csv('allrounder_2014.csv', header=T)
allrounder2015.df<-read.csv('allrounder_2015.csv', header=T)
allrounder2016.df<-read.csv('allrounder_2016.csv', header=T)
allrounder2017.df<-read.csv('allrounder_2017.csv', header=T)

odi_batting2014.df<-read.csv('odi_batsman_2014.csv', header=T)
odi_batting2015.df<-read.csv('odi_batsman_2015.csv', header=T)
odi_batting2016.df<-read.csv('odi_batsman_2016.csv', header=T)
odi_batting2017.df<-read.csv('odi_batsman_2017.csv', header=T)


odi_bowling2014.df<-read.csv('odi_baller_2014.csv', header=T)
odi_bowling2015.df<-read.csv('odi_baller_2015.csv', header=T)
odi_bowling2016.df<-read.csv('odi_baller_2016.csv', header=T)
odi_bowling2017.df<-read.csv('odi_baller_2017.csv', header=T)


odi_allrounder2014.df<-read.csv('odi_allrounder_2014.csv', header=T)
odi_allrounder2015.df<-read.csv('odi_allrounder_2015.csv', header=T)
odi_allrounder2016.df<-read.csv('odi_allrounder_2016.csv', header=T)
odi_allrounder2017.df<-read.csv('odi_allrounder_2017.csv', header=T)
```

# Setting Training data
```{r 3}
batting_train<-rbind(batting2014.df,batting2015.df,batting2016.df,odi_batting2014.df,odi_batting2015.df,odi_batting2016.df)
bowling_train<-rbind(bowling2014.df,bowling2015.df,bowling2016.df,odi_bowling2014.df,odi_bowling2015.df,odi_bowling2016.df)
allrounder_train<-rbind(allrounder2014.df,allrounder2015.df,allrounder2016.df,odi_allrounder2014.df,odi_allrounder2015.df,odi_allrounder2016.df)

#Remove Na rows
batting_train<-na.omit(batting_train)
bowling_train<-na.omit(bowling_train)
allrounder_train<-na.omit(allrounder_train)

batting_train<- batting_train[ ,-c(1,2,11) ]
bowling_train<-bowling_train[ ,-c(1,2,8,12) ]
allrounder_train<-allrounder_train[ ,-c(1,2,7,12) ]


```

# Setting Testing data
```{r 4}

batting_test<-rbind(batting2017.df,odi_batting2017.df)
bowling_test<-rbind(bowling2017.df,odi_bowling2017.df)
allrounder_test<-rbind(allrounder2017.df,odi_allrounder2017.df)
#Remove Na rows
batting_test<-na.omit(batting_test)
bowling_test<-na.omit(bowling_test)
allrounder_test<-na.omit(allrounder_test)

batting_test<- batting_test[ ,-c(1,2,11) ]
bowling_test<-bowling_test[ ,-c(1,2,8,12) ]
allrounder_test<-allrounder_test[ ,-c(1,2,7,12) ]
```

##Model 1: Simple Linear Regression
```{r 41}
batting_regress<- lm(Price~.,data=batting_train)
bowling_regress<- lm(Price~.,data=bowling_train)
allrounder_regress <- lm(Price~.,data=allrounder_train)
```

# Regression summary
```{r 5}
#Batting
par(mfrow=c(2,2))
summary(batting_regress)
plot(batting_regress)
```
```{r 6}
#Bowling
summary(bowling_regress)
plot(bowling_regress)
```
```{r 7}
#Allrounder
summary(allrounder_regress)
plot(bowling_regress)
```

```{r 8}
#Simple linear regression Prediction
batting_regress_predict<-predict(batting_regress,batting_test[,-c(9)])
bowling_regress_predict<-predict(bowling_regress,bowling_test[,-c(9)])
allrounder_regress_predict<-predict(allrounder_regress,allrounder_test[,-c(9)])

```
#RMSE
```{r 9}
batting_SLR_RMSE<-sqrt(sum((batting_regress_predict- batting_test$Price)^2 ))
bowling_SLR_RMSE<-sqrt(sum((bowling_regress_predict- bowling_test$Price)^2 ))
allrounder_SLR_RMSE<-sqrt(sum((allrounder_regress_predict- allrounder_test$Price)^2 ))
batting_SLR_RMSE
bowling_SLR_RMSE
allrounder_SLR_RMSE
```

##Model 2: Ridge Regression
```{r 10}
# X variables
batting_x<- batting_train[ ,-c(12) ]
bowling_x<-bowling_train[ ,-c(13) ]
allrounder_x<-allrounder_train[ ,-c(13)]

set.seed(123)
batting_ridge<- cv.glmnet (as.matrix(batting_x),batting_train$Price,lambda = 10^seq(4,-1,-0.1),alpha=0 )
bowling_ridge<- cv.glmnet(as.matrix(bowling_x),bowling_train$Price,lambda = 10^seq(4,-1,-0.1),alpha=0)
allrounder_ridge <- cv.glmnet(as.matrix(allrounder_x),allrounder_train$Price,lambda = 10^seq(6,-1,-0.1),alpha=0)
```

# Ridge Lambda selection and fitting
```{r 11}
#Batting
par(mfrow=c(1,1))
plot(batting_ridge)
best_batting_lambda<-batting_ridge$lambda.1se
best_batting_lambda
batting_ridge_coeff<-batting_ridge$glmnet.fit$beta[batting_ridge$glmnet.fit$lambda==best_batting_lambda]
batting_ridge_r2 <- batting_ridge$glmnet.fit$dev.ratio[which(batting_ridge$glmnet.fit$lambda == best_batting_lambda)]
batting_ridge_r2
```
```{r 12}
#Bowling
plot(bowling_ridge)
best_bowling_lambda<-bowling_ridge$lambda.1se
best_bowling_lambda
bowling_ridge_coeff<-bowling_ridge$glmnet.fit$beta[bowling_ridge$glmnet.fit$lambda==best_bowling_lambda]
bowling_ridge_r2 <- bowling_ridge$glmnet.fit$dev.ratio[which(bowling_ridge$glmnet.fit$lambda == best_bowling_lambda)]
bowling_ridge_r2
```
```{r 13}
#Allrounder
plot(allrounder_ridge)
best_allrounder_lambda<-allrounder_ridge$lambda.1se
best_allrounder_lambda
allrounder_ridge_coeff<-allrounder_ridge$glmnet.fit$beta[allrounder_ridge$glmnet.fit$lambda==best_allrounder_lambda]
allrounder_ridge_r2 <- allrounder_ridge$glmnet.fit$dev.ratio[which(allrounder_ridge$glmnet.fit$lambda == best_allrounder_lambda)]
allrounder_ridge_r2
```

# Prediction for Ridge Regression
```{r 14}
batting_ridge_predict<-predict(batting_ridge, as.matrix(batting_x),s=c("lambda.1se"))
bowling_ridge_predict<-predict(bowling_ridge, as.matrix(bowling_x),s=c("lambda.1se"))
allrounder_ridge_predict<-predict(allrounder_ridge, as.matrix(allrounder_x),s=c("lambda.1se"))
```

#Model 3: Lasso regression

```{r 15}

set.seed(123)
batting_lasso<- cv.glmnet (as.matrix(batting_x),batting_train$Price,lambda = 10^seq(4,-1,-0.1),alpha=1 )
bowling_lasso<- cv.glmnet(as.matrix(bowling_x),bowling_train$Price,lambda = 10^seq(4,-1,-0.1),alpha=1)
allrounder_lasso <- cv.glmnet(as.matrix(allrounder_x),allrounder_train$Price,lambda = 10^seq(6,-1,-0.1),alpha=1)
```

# Lasso Lambda selection and fitting
```{r 16}
#Batting
par(mfrow=c(1,1))
plot(batting_lasso)
best_batting_lambda<-batting_lasso$lambda.1se
best_batting_lambda
batting_lasso_coeff<-batting_lasso$glmnet.fit$beta[batting_lasso$glmnet.fit$lambda==best_batting_lambda]
batting_lasso_r2 <- batting_lasso$glmnet.fit$dev.ratio[which(batting_lasso$glmnet.fit$lambda == best_batting_lambda)]
batting_lasso_r2
```
```{r 17}
#Bowling
plot(bowling_lasso)
best_bowling_lambda<-bowling_lasso$lambda.1se
best_bowling_lambda
bowling_lasso_coeff<-bowling_lasso$glmnet.fit$beta[bowling_lasso$glmnet.fit$lambda==best_bowling_lambda]
bowling_lasso_r2 <- bowling_lasso$glmnet.fit$dev.ratio[which(bowling_lasso$glmnet.fit$lambda == best_bowling_lambda)]
bowling_lasso_r2
```
```{r 18}
#Allrounder
plot(allrounder_lasso)
best_allrounder_lambda<-allrounder_lasso$lambda.1se
best_allrounder_lambda
allrounder_lasso_coeff<-allrounder_lasso$glmnet.fit$beta[allrounder_lasso$glmnet.fit$lambda==best_allrounder_lambda]
allrounder_lasso_r2 <- allrounder_lasso$glmnet.fit$dev.ratio[which(allrounder_lasso$glmnet.fit$lambda == best_allrounder_lambda)]
allrounder_lasso_r2
```

# Prediction for Lasso Regression
```{r 19}
batting_lasso_predict<-predict(batting_lasso, as.matrix(batting_x),s=c("lambda.1se"))
bowling_lasso_predict<-predict(bowling_lasso, as.matrix(bowling_x),s=c("lambda.1se"))
allrounder_lasso_predict<-predict(allrounder_lasso, as.matrix(allrounder_x),s=c("lambda.1se"))
```


#R Square computation
```{r 191}
rsquare <- function(true, predicted) {
  sse <- sum((predicted - true)^2)
  sst <- sum((true - mean(true))^2)
  rsq <- 1 - sse / sst

  # For this post, impose floor...
  if (rsq < 0) rsq <- 0

  return (rsq)
}
```

# R Square Comparison

```{r 20}


batting_regress_r2<-summary(batting_regress)$r.squared
bowling_regress_r2<-summary(bowling_regress)$r.squared
allrounder_regress_r2<-summary(allrounder_regress)$r.squared
batting_regress_r2
bowling_regress_r2
allrounder_regress_r2

batting_ridge_r2<-rsquare(batting_train$Price,predict(batting_ridge, as.matrix(batting_x),s=c("lambda.1se")))
bowling_ridge_r2<-rsquare(bowling_train$Price,predict(bowling_ridge, as.matrix(bowling_x),s=c("lambda.1se")))
allrounder_ridge_r2<-rsquare(allrounder_train$Price,predict(allrounder_ridge, as.matrix(allrounder_x),s=c("lambda.1se")))
batting_ridge_r2
bowling_ridge_r2
allrounder_ridge_r2

batting_lasso_r2<-rsquare(batting_train$Price,predict(batting_lasso, as.matrix(batting_x),s=c("lambda.1se")))
bowling_lasso_r2<-rsquare(bowling_train$Price,predict(bowling_lasso, as.matrix(bowling_x),s=c("lambda.1se")))
allrounder_lasso_r2<-rsquare(allrounder_train$Price,predict(allrounder_lasso, as.matrix(allrounder_x),s=c("lambda.1se")))
batting_lasso_r2
bowling_lasso_r2
allrounder_lasso_r2


```



